---
title: "Lab2"
author: "Samson Olivero"
date: "2/4/2021"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tasks

## Task 1

Get working directory

```{r}
getwd()
```

## Task 2

Read data into R

```{r}
epagas = read.csv("EPAGAS.csv")
```

Get the first six lines of data

```{r}
head(epagas)
```

## Task 3

Make the Object MPG vector 

```{r}
mpg = epagas$MPG #Creates a vector of the object MPG in epagas file
```

Transform MPG to z and verify

```{r}
z = (mpg - mean(mpg)) / sd(mpg) #formula found in word file 
head(z)
#paste0() function in R Language is used to concatenate all elements without separator.
#round answer to the 4th decimal point
#Get rid of quotation marks
print(paste0("Mean = ", round(mean(z), digits = 4)), quote = FALSE)
#variance is a measure of how spread out a data set is. It is calculated as the average squared deviation of each number from the mean of a data set
print(paste0("Variance = ", var(z)), quote = FALSE)
#Standard deviation is the measure of dispersion of a set of data from its mean. It measures the    absolute variability of a distribution
print(paste0("Standard Deviation = ", sd(z)), quote = FALSE)
```


Finding possible MPG outliers using z

```{r}
outliers = mpg[abs(z) >= 2 & abs(z) <= 3]
outliers
```


Finding defined MPG outliers using z

```{r}
defOutliers = mpg[abs(z) > 3]
defOutliers
```


Using Lattice Package construct colored dotplot

```{r}
#Red = Outlier; Blue= Possible Outlier
mpgcol = 
  ifelse(
  abs(z) > 3, "Red",
  ifelse(abs(z) >= 2 & abs(z) <= 3, "Blue",
    "Black"
    )
  )
library(lattice)
dotplot(mpg, col = mpgcol, main = "Dotplot of MPG Values")
```


## Task 4

Make a boxplot of mpg variables

Use Chebyshev's theorem to predict the proportion of data within 2 standard deviations of the mean of mpg data

Calculate Exact Proportion within 2 standard deviations of the mean

Does Chebyshev Agree with the data ?

Empirical Rule: what proportion of the data should be within 2 standard deviations of the mean

How well does it correspond ?

Is the Empircal Rule Valid ? Why ?



