---
title: "Lab2"
author: "Samson Olivero"
date: "2/4/2021"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tasks

## Task 1

Get working directory

```{r}
getwd()
```

## Task 2

Read data into R

```{r}
epagas = read.csv("EPAGAS.csv")
```

Get the first six lines of data

```{r}
head(epagas)
```

## Task 3

Make the Object MPG

Transform MPG to z and verify

Finding possible MPG outliers using z

Finding defined MPG outliers using z

Using Lattice Package construct colored dotplot

```{r}
library(lattice)
#dotplot(mpg, col=mpgcol)
```


## Task 4

Make a boxplot of mpg variables

Use Chebyshev's theorem to predict the proportion of data within 2 standard deviations of the mean of mpg data

Calculate Exact Proportion within 2 standard deviations of the mean

Does Chebyshev Agree with the data ?

Empirical Rule: what proportion of the data should be within 2 standard deviations of the mean

How well does it correspond ?

Is the Empircal Rule Valid ? Why ?



